---
title: "Learning Diary"
author: "Xinyuan Zhang"
date: "17/06/2021"
output: pdf_document
---

## General notes

this is some notes for everyday work and remember some questions. Should update daily. May add sections. May want to add to overleaf, but that's maybe for future. Each note/question starts with a dot point now, but maybe change format later. 

## Notes to self

### 16/07

* in codes, I guess it's from Menzel's notation, $\beta$ denotes for model reated to f, while $\Gamma$ denotes for model related to g, and when cts case, the spline smoothing function. Or maybe I'm wrong? *need confirm*

* readed the **rpm.loglik.cts.R**, that's initial def of loglikfun_cts, eqfun_cts, gloglikfun_cts and jeqfun_cts. Straightforward, but need to understand loglik & eqcond defined in Rcpp, which may need change too for cts var

* Now read **loglik.cpp**

### 17/07

* mostly confirmed: loglik.cpp defines loglik as a function

* Rcpp gallery: good reference https://gallery.rcpp.org/

* **loglik.cpp** : it's defining W as U+V+Geo, which is about S; I may want to start w/ sth simpler though

* I basically understand **loglik** defined in **loglik.cpp**, that basically expresses (12) and return pseudo liglik value; tech details qeustioning though

* Now read **eqcond.cpp**. But since we have **eqcondR.R**, start with this first

* **eqcondR.R** : output is constraint from (8'), should be 0 if met, because the eqM and eqW are RHS - LHS = 0. 

* some details about **eqcondR.R** : 

  * what are the counts doing? I assume it will be clear once I can work with data. 
  
  * What's the meaning of S, W, M, nXu, nZu? I need to know more clearly. Now I just have general idea 

  * Once I know the things above I'll get what beta contains. 

* Now, read **eqcond.cpp **, which half of it is same as **eqcondR.R**, but contains how to use the constraint calculated

* some details about **eqcond.cpp** : 

  * pfp = pfps - exp(-GammaW) is actually the constraint from (8'), again, this is RHS - LHS should be 0
  
  * If constraints==0 that's constraint satisfied: pfp[more index] = pmf_something_with_index_woman - f(*,z)

* Now, start to read **rpm_cts.R**

### 18/07

* I may have to know more about control.

* some details about **rpm_cts.R** : 

  * line 270&271, gw and gm are log(no. of women or men / total)
  
* The working pattern according to  **rpm_cts.R**:

  * data cleaning, weighting, get proportion of men & women, etc
  
  * create model matrix; codes seems like working on **distance** for S
  
  * need confirm: is alpha is the spline function alpha? 
  
  * set initial theta. now quite understand the code
  
  * Proper algorithm begins
  
  * insert value from initial theta to containers called beta! lower bound & upper bound
  
  * use eqfun_cts (should use functions defined in Rcpp as eqcond) to calculate a test_eqcond
  
  * call **nloptr::nloptr** to calculate a out / out.text. Put a some other values in; like hessian, etc
  
  * out$solution is th_hat. Use this value into eqfun_cts instead of initial theta
  
  * use a rpm.hessian_cts function somewhere in the process
  
  * Put a whole bunch of other things into the **out** stuff
  
  * then do bootstrap

* **rpm.loglik.cts.R** use **numDeriv::grad** for gloglikfun_cts and **nloptr::nl.jacobian** (same as **nl.grad** as the guide suggests) for jeqfun_cts

* I think now it's a good time to temporarily stop reading through codes and start reading or working on some math, with checking **nloptr** and **cubature** in mind. 

* Also, may want to read through **control.rpm.R** later

* What's needed: numerical quadrature & cubature package; spline smooth model & spline2 package.

* Now, read something about spline smooth

### 19/07

* Reexpress of g: $g(x,*)=\sum_{k=1}^{K}{\alpha^W_k b_k(x)}$

* vedio : intro to smoothing splines https://www.youtube.com/watch?v=queK1reC-ac

  * a smoothing spline is a spline designed to balance fit with smoothness
  
* intro of Bernstein Polynomials https://www.youtube.com/watch?v=AL0vcsLlYp4

  * Definition of *Berndtein basis Poly*: $b_{i,n}(x)= \binom{n}{i}x^i(1-x)^{n-i}$
  
  * Linear combination of Bernstein basis poly of degree n: $B_k(x)=\sum^K_{k=0}\alpha_kb_{k,n}(x)$

*So this is question part: need to confirm if the reexpression of g is just the linear combinatino of B poly

* And yes, now the question is about estimating $\alpha_k$

* article of package **splines2**: https://wwenjie.org/splines2/articles/splines2-intro

  * here at last, generalized Bernstein poly basis of degree n over [a,b] is : $B_i^n(x)=\frac{1}{(b-a)^n}\binom{n}{i}(x-a)^i(b-x)^{n-i}$, i from 0 to n, which reduce to normal when a=0 and b=1. 

* Now, leave the math details first, and check **cubature** package and **numerical quadrature**

* Honestly not quite understand these two. Assumption for now: **cubature** is for integration instead of summation when calculating eqcond. need confirm. If that's the case then it's much clearer. 

* Summation now used for eqcond and loglik, but integration is most likely for eqcond since (12) involves summation? need confirm

### 20/07

* delete after starting: may want to look into **nloptr** and **cubature**. 

* Rcpp reference guide: https://dirk.eddelbuettel.com/code/rcpp/Rcpp-quickref.pdf

* github & R studio intro bookdown https://happygitwithr.com/rstudio-git-github.html , video https://www.youtube.com/watch?v=QLFc9gw_Hfs&t=786s , https://www.youtube.com/watch?v=megZYkCLMA4



### 22/07

* Read through some comments in overleaf doc. Help to understand Gamma_W maybe? $\frac{\exp{\lbrace W(x,z) \rbrace} w(x) m(z)}{(1+ \Gamma_w(x))(1+ \Gamma_m(z))}$

$$
f(x,z) = \frac{\exp{\lbrace W(x,z) \rbrace} w(x) m(z)}{(1+ %\Gamma_w(x))(1+ \Gamma_m(z))} = 
         2e^{W(x,z|\beta)} f(x,*) f(*,z)~~~~~~~~\forall x, z  %\label{eq:fxz}
$$
$$
2\frac{N_m}{N}e^{-g(x,*)}+\frac{N_w-N_m}{N} = \int{\frac{e^{W(x,s)+g(*,s)}m(s)}{1+e^{g(*,s)}}ds}~~~\forall~x \\
2\frac{N_w}{N}e^{-g(*,z)}-\frac{N_w-N_m}{N} = \int{\frac{e^{W(x,s)+g(s,*)}w(s)}{1+e^{g(s,*)}}ds}~~~\forall~z
$$

$$
e^{-g(x,*)} = \int{\frac{e^{W(x,s)+g(*,s)}m(s)}{1+e^{g(*,s)}}ds}~~~\forall~x\\
e^{- \sum_{k=1}^{K}{\alpha^W_k b_k(x)}} = \int{\frac{e^{W(x,s)+ \sum_{k=1}^{K}{\alpha^W_k b_k(s)}}m(s)}{1+e^{ \sum_{k=1}^{K}{\alpha^W_k b_k(s)}}}ds}~~~\forall~x
$$
* check about how the re expression of $g(*,z)$ different from g(x,*).

### 23/07

* Tomorrow should check the recordings

### 25/07

* sorry for not being very productive this week

* should do better today. I still have couple of hours! Don't waste it!

* first thing: try to do a simple g function based on **splines2** package

```{r}
library(splines2)
x1 <- seq.int(0, 1, 0.1)
x2 <- seq.int(- 1, 1, 0.01)
bpMat1 <- bernsteinPoly(x1, degree = 4, intercept = TRUE)
bpMat2 <- bernsteinPoly(x1, degree = 4, intercept = T)
dim(bpMat1)
length(x1)
```
```{r}
x1 <- seq.int(0, 1, 0.1)
x2 <- seq.int(- 1, 1, 0.01)
bpMat1 <- bernsteinPoly(x1, degree = 4, intercept = F)
bpMat1
```

* Coding questions: 

  * What's K? Is it unique x n z respectively? 

* Tomorrow: do integration with **cubature**

### 31/07

* fix part of loglik.R.cts and eqcond.R.cts

* tomorrow: listen to recording of last meeting, fix loglik.R.cts the left part of counts; should I add indicators?

* probably have to get t know nloptr for estimating

* try to get gradient and Jacobin to work by tomorrow

### 01/08

* fixed many errors in eqcond function; remember to update tomorrow to github

* big problem: cannot do matching inside function for cubature! sol: find closest value from X

* return lots of NA

* solved by changing matching methods; but still does not seem satisfying

* tomorrow: maybe talk to Mark about the problem with basis matrix; the values returned are just too small to be used; results will be too heavily based on alpha, constraint closer to 0 when smaller parameters assumed, which should not be the case?

* tomorrow: fix matching for loglik, try to run it till it works

### 02/08

* Now we confirmed: eqcond, eqcond.5p, loglik all works fine

* loglikfun_cts, eqfun_cts works; gradient and jacobian take too long;

* wrote gloglik.cts mathematically; need check and run

* not sure how to write jacobian though

### 03/08

* after a night nl.jacobian doesn't work

* gloglik_fun works with written gradient; need check math

* do we need match for cubature integration? or should we put it to poly?


### 11/08

* change **splines2** into **prama::bernsteinb** so don't have to match in function for eqcond when doing integration; 

* checked manually, not sure why splines2 gives the wrong number result

* problem with **splines2**: gives wrong value (make not be a big deal since values are really small), and changes with diff length of X

* but **prama** cannot give a matrix, better one input at a time; leave it like this for now, fix later

* new **eqcond.R** function works fine!

* tomorrow: listen to rest of recordings; should start fix **loglik**; try to finish by Saturday

### 12/08

* **eqcond.jacobian** works! result in a 10*202 matrix

### 13/08

* fix **loglik.R** with separated data for paired and single

* now numeric gradient works!

* tomorrow should try to do optimization with 200



\

## Questions along reading the codes


* syntax questions; for numvector & intvector, when to use [] and ()? 

* in *loglik.cpp*, the llik part, what's gm and gw (both is double)? Still not sure. Is GammaM(k)+gm representing $g(*,z)$? *A*: I think yes based on **eqcondR.R**. 


* in **eqcond.cpp**, what's pfp at last is doing? What's pmf? Is pfp before the constraint that should be zero ? 

* What's all the wieighting doing? I think it's mentioned in the paper. need check



* What's **gloglik.cpp** last part about partial g(x,*) doing? what's the gf it creates?




